Probabilistic Inference of NYC Congestion from Taxi Data

### 2017-12-09 reports/paper/paper.tex
Add notebooks and paper background.

\usepackage{times}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{natbib}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{lipsum}
\usepackage{todonotes,tikz}
\usepackage{cite}
\usepackage{natbib}

@@

\begin{algorithm}
  \begin{algorithmic}
    \STATE{\COMMENT{Optimizing using Adam in PyTorch}}

@@

    Update $\bm W$.
    } \ENDFOR
  \end{algorithmic}
  \caption{Training algorithm for softmax path selection}
  \label{alg:main_alg}
\end{algorithm}
}

@@

@@

In addition to adding columns, the original dataset was filtered. Geographic
coordinates were\ldots

>>>>

\usepackage{times}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{natbib}
\usepackage{algorithm}
\usepackage{algorithmic,inconsolata}
\usepackage{lipsum}
\usepackage{todonotes,tikz}
\usepackage{cite}
\usepackage{natbib}

@@

}
\begin{algorithm}
  \begin{algorithmic}
    \STATE{\COMMENT{Optimizing using \texttt{torch.optim.Adam}}}

@@

    Update $\bm W$.
    } \ENDFOR
  \end{algorithmic}
  \caption{Training algorithm for softmax path selection}
  \label{alg:main_alg}
\end{algorithm}

@@

\subsection{Data Processing}

@@

In addition to adding columns, the original dataset was filtered. Only trips
whose start and locations fell inside a rectangle that bounds Manhattan were
considered. As a preliminary filtering, trips with duration less than 0 or
greater than one day were removed. Upon training models, this was further
culled to trips between two minutes and two hours in length.

Principal component analysis was run on all de-meaned and
standardized geographic coordinates (both start
and end locations) from the filtered data. The coordinates were thus
transformed into the first and second principal components. For the discrete
grid models, these coordinates were further discretized into a $20 \times 70$
grid.

The data-set was saved as a whole, but also partioned into $(\text{day of week},
\text{hour})$ groups, using the one-hot encoded vectors. Some models were
trained on the entire data-set (e.g.\ Neural Network), and some were trained
individually on the partioned data-sets (e.g.\ Structural Models).

\subsection{Baselines}

Three baselines were calculated. The first was a linear regression predicting
trip duration from trip distance (calculated using the Pythagorean Theorem and
the start and end coordinates), and trained on a partitioned data set.

The second was a neural network trained on the
entire dataset taking as input the start and end coordinates as well as the day
of week and hour one-hot encoded vectors. The neural network had one hidden
layer with 50 neurons and a tanh activation function.

The second baseline was constructed by querying the Google Maps Distance Matrix
API\footnote{\url{https://developers.google.com/maps/documentation/distance-matrix/}}.
While there was no need to train this baseline, there was a rate limit
constraint on how many test data points it could be tested on. We assess its
performance on 2,500 data points in a representative partioned test data-set.

\subsection{Training}

Training implementations varied from model to model. In training the neural
network baseline, SGD was used with a learning rate of $10^{-7}$ run for 12
epochs on the full data set. This took between 30 minutes and an hour. The
uniform paths structural model was similarly trained with SGD, with a learning
rate of $10^{-4}$.

The uniform model was only trained on a subset of the
dataset corresponding to all trips on a given day of the week and on a given
hour. Convergence appeared after roughly two epochs, and the model was run for 5
epochs taking just under ten minutes.

The softmax structural model was trained
using the \texttt{Adam} optimizer with learning rate $10^{-2}$ for 10 epochs,
taking approximately an hour.
% TODO: jchen confirm this for me:
